---
title: 'Class 6a: Blake et al. (2015) exercise'
author: "Mike Frank"
date: "February 9, 2016"
output: 
  html_document:
    toc: true
---

# Intro

This is an in-class exercise exploring Blake et al. (2015), [Ontogeny of fairness in seven societies](http://www.nature.com/nature/journal/v528/n7581/full/nature15703.html).

Please explore these data together (without looking at the analyses supplied by the authors). 

The overall goal is to understand the degree to which the data support the authors' hypotheses. 

```{r}
rm(list=ls())
library(ggplot2)
library(tidyr)
library(dplyr)
library(bootstrap)

theme_set(theme_bw())
sem <- function(x) {sd(x, na.rm=TRUE) / sqrt(length(x))}
ci95 <- function(x) {sem(x) * 1.96}

theta <- function(x,xdata,na.rm=T) {mean(xdata[x],na.rm=na.rm)}
ci.low <- function(x,na.rm=T) {
  mean(x,na.rm=na.rm) - quantile(bootstrap(1:length(x),1000,theta,x,na.rm=na.rm)$thetastar,.025,na.rm=na.rm)}
ci.high <- function(x,na.rm=T) {
  quantile(bootstrap(1:length(x),1000,theta,x,na.rm=na.rm)$thetastar,.975,na.rm=na.rm) - mean(x,na.rm=na.rm)}

```

# Data Prep

First read in the data, as distributed by the journal. 

```{r}
d <- read.csv("data/Ontogeny_fairness_seven_societies_data.csv", 
                header = TRUE, na.strings = c("NA", "."))
```

Do some preprocessing, taken directly from the supplemental material. 

```{r}
facVars <- c("eq.uneq", "value", "decision")
d[, facVars] <- lapply(d[, facVars], factor)
d$trial.number <- as.numeric(gsub(".(\\d+)", "\\1", d$trial))
```

Rename things so that they are easy to deal with. 

```{r}
d$trial_type <- factor(d$eq.uneq, 
                       levels = c("E","U"), 
                       labels = c("Equal","Unequal"))
d$condition <- factor(d$condition,
                      levels = c("AI","DI"), 
                      labels = c("Advantageous","Disadvantageous"))
```

# Variable exploration

Describe the dataset graphically in ways that are useful for you to get a handle on the data collection effort. 

```{r}
demo <- d %>%
  mutate(age = floor(actor.age.years)) %>%
  group_by(country, age, actor.id) %>%
  summarise(n = n()) %>%
  summarise(n = n())

ggplot(demo, aes(x = age, y = n)) + 
  geom_bar(stat = "identity") + 
  facet_grid(. ~ country)
```

# Hypothesis-related exploration

In this second, explore the authors' hypotheses related to advantageous and inadvantageous inequity aversion. Create 1 - 3 pictures that describe the support (or lack of it) for this hypothesis. 

Try just plotting the un-adjusted curves. 

```{r}
ms <- d %>%
  filter(!is.na(eq.uneq)) %>%
  mutate(age = floor(actor.age.years), 
         decision = decision == "reject") %>%
  group_by(country, trial_type, condition, age, actor.id) %>%
  summarise(decision = mean(decision, na.rm=TRUE)) %>%
  summarise(mean = mean(decision, na.rm=TRUE), 
            ci_lower = mean(decision) - ci95(decision),
            ci_upper = mean(decision) + ci95(decision),
            n = n())

ggplot(ms, aes(x = age, y = mean, col = country)) + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
                  position = position_dodge(width = .1)) + 
  geom_point(aes(size = n)) +
  facet_grid(condition ~ trial_type) + 
  geom_smooth(method="lm", se = FALSE) + 
  ylab("Proportion offers rejected") +
  xlab("Age (years)") +
  ylim(c(0,1)) 
```

Now rebin into 3-year bins.

```{r}
ms <- d %>%
  filter(!is.na(eq.uneq)) %>%
  mutate(age = floor(actor.age.years/2)*2, 
         decision = decision == "reject") %>%
  group_by(country, trial_type, condition, age, actor.id) %>%
  summarise(decision = mean(decision, na.rm=TRUE)) %>%
  summarise(mean = mean(decision, na.rm=TRUE), 
            ci_lower = mean(decision) - ci95(decision),
            ci_upper = mean(decision) + ci95(decision),
            n = n())

ggplot(ms, aes(x = age, y = mean, col = country)) + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
                  position = position_dodge(width = .1)) + 
  geom_point(aes(size = n)) +
  facet_grid(condition ~ trial_type) + 
  geom_smooth(method="lm", se = FALSE) + 
  ylab("Proportion offers rejected") +
  xlab("Age (years)") +
  ylim(c(0,1)) 
```

Break this down by country. I like this plot best so far. 

```{r}
ggplot(ms, aes(x = age, y = mean, col = condition)) + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
                  position = position_dodge(width = .1)) + 
  geom_point(aes(size = n)) +
  facet_grid(trial_type ~ country) + 
  geom_smooth(method="lm", se = FALSE, span = 2) + 
  ylab("Proportion offers rejected") +
  xlab("Age (years)") +
  ylim(c(0,1)) 
```

But maybe we can do better. Let's:

+ bootstrap CIs
+ sort the plots
+ add better smoother (quadratic)
+ use lines instead of points
+ better palette

```{r}
ms <- d %>%
  filter(!is.na(eq.uneq)) %>%
  mutate(age = floor(actor.age.years/2)*2, 
         decision = decision == "reject") %>%
  group_by(country, trial_type, condition, age, actor.id) %>%
  summarise(decision = mean(decision, na.rm=TRUE)) %>%
  summarise(mean = mean(decision, na.rm=TRUE), 
            ci_lower = ci.low(decision),
            ci_upper = ci.high(decision),
            n = n())

ms$country <- factor(ms$country, 
                     levels = c("Uganda","US","Canada",
                                "Senegal","India","Peru","Mexico"))

ggplot(ms, aes(x = age, y = mean, col = condition)) + 
  geom_linerange(aes(ymin = mean - ci_lower, 
                     ymax = mean + ci_upper), 
                  position = position_dodge(width = .5)) + 
  geom_point() +
  facet_grid(trial_type ~ country) + 
  geom_smooth(method="lm", se = FALSE, formula = y ~ I(x^2)) + 
  ylab("Proportion offers rejected") +
  xlab("Age (years)") +
  ylim(c(0,1)) + 
  langcog::scale_colour_solarized()
```
