---
title: 'Class 6a: Blake et al. (2015) exercise'
author: "Mike Frank"
date: "February 9, 2016"
output: 
  html_document:
    toc: true
---

# Intro

This is an in-class exercise exploring Blake et al. (2015), [Ontogeny of fairness in seven societies](http://www.nature.com/nature/journal/v528/n7581/full/nature15703.html).

Please explore these data together (without looking at the analyses supplied by the authors). 

The overall goal is to understand the degree to which the data support the authors' hypotheses. 

```{r}
rm(list=ls())
library(ggplot2)
library(tidyr)
library(dplyr)
library(lme4)

theme_set(theme_bw())
sem <- function(x) {sd(x, na.rm=TRUE) / sqrt(length(x))}
ci95 <- function(x) {sem(x) * 1.96}
```

# Data Prep

First read in the data, as distributed by the journal. 

```{r}
d <- read.csv("data/Ontogeny_fairness_seven_societies_data.csv", 
                header = TRUE, na.strings = c("NA", "."))
```

Do some preprocessing, taken directly from the supplemental material. 

```{r}
facVars <- c("eq.uneq", "value", "decision")
d[, facVars] <- lapply(d[, facVars], factor)
d$trial.number <- as.numeric(gsub(".(\\d+)", "\\1", d$trial))
```

Rename things so that they are easy to deal with. 

```{r}
```

# Variable exploration

Describe the dataset graphically in ways that are useful for you to get a handle on the data collection effort. 

```{r}
```

Make sure you understand what the design was: how many trials per participant, what was between- and within-subject, etc. 

# Hypothesis-related exploration

In this second, explore the authors' hypotheses related to advantageous and inadvantageous inequity aversion. Create 1 - 3 pictures that describe the support (or lack of it) for this hypothesis. 

```{r}
```


```{r}
```

