---
title: 'Psych 254 W15 PS #1'
author: "Mike Frank"
date: "February 16, 2017"
output: html_document
---

This is problem set #1, in which we hope you will practice the packages tidyr and dplyr. There are some great cheat sheets from [RStudio](http://www.rstudio.com/resources/cheatsheets/).

The data set
------------

This data set comes from a replication of [Janiszewski and Uy (2008)](http://dx.doi.org/10.1111/j.1467-9280.2008.02057.x), who investigated whether the precision of the anchor for a price influences the amount of adjustment.

In the data frame, the `Input.condition` variable represents the experimental condition (under the rounded anchor, the rounded anchor, over the rounded anchor). `Input.price1`, `Input.price2`, and `Input.price3` are the anchors for the `Answer.dog_cost`, `Answer.plasma_cost`, and `Answer.sushi_cost` items.

Preliminaries
-------------

I pretty much always clear the workspace and load the same basic helper functions before starting an analysis.

```{r prelims}
rm(list=ls())
source("helper/useful.R")
library(stringr)
library(readr)
```

Note that I'm using a "relative" path (the "helper") rather than an absolute path (e.g. "/Users/mcfrank/code/projects/etc..."). The relative path means that someone else can run your code by changing to the right directory, while the absolute path will force someone else to make trivial changes every time they want to run it.

Part 1: Data cleaning
---------------------

The first part of this exercise actually just consists of getting the data in a format usable for analysis. This is not trivial. Let's try it:

```{r data1}
d <- read.csv("data/janiszewski_rep_exercise.csv")
```

Fine, right? Why can't we go forward with the analysis?

HINT: try computing some summary statistics for the different items. Also, are there any participants that did the task more than once?

> ANSWER-START
There is non-uniformity in the Input and Answer formats ("five hundred", an empty string, "5,012", "7,75"). Additionally, two participants did the task more than once: A1VYX5VKZ0CTWU and A5IUO2T5MZQUZ.
> ANSWER-END

Fix the data file *programmatically*, i.e., write code that transforms the unclean data frame into a clean data frame.

> ANSWER-START

```{r}
d <- read_csv("data/janiszewski_rep_exercise.csv", na = c("", "NA"))

summary(as.factor(d$Answer.sushi_cost)) # a shortcut to see unique values
summary(as.factor(d$Answer.dog_cost)) 
summary(as.factor(d$Answer.plasma_cost)) 

d$Answer.sushi_cost <- str_replace_all(d$Answer.sushi_cost, "ehight", "8")
d$Answer.sushi_cost <- as.numeric(str_replace_all(d$Answer.sushi_cost, ",", "."))

d$Answer.dog_cost <- str_replace_all(d$Answer.dog_cost, "five hundred", "500")
d$Answer.dog_cost <- as.numeric(str_replace_all(d$Answer.dog_cost, ",", ""))

d$Answer.plasma_cost <- as.numeric(d$Answer.plasma_cost)

```
> ANSWER-END

Part 2: Making these data tidy
------------------------------

Now let's start with the cleaned data, so that we are all beginning from the same place.

```{r data2}
d <- read.csv("data/janiszewski_rep_cleaned.csv")
```

This data frame is in *wide* format - that means that each row is a participant and there are multiple observations per participant. This data is not *tidy*.

To make this data tidy, we'll do some cleanup. First, remove the columns you don't need, using the verb `select`.

HINT: `?select` and the examples of helper functions will help you be efficient.

```{r select}
d.tidy <- select(...)
```

> ANSWER-START

```{r select}
d.tidy <- select(d, WorkerId, Input.condition, starts_with("Answer"))
```
> ANSWER-END

Try renaming some variables using `rename`. A good naming scheme is:

* consistent with case
* consistent with "." or "_" ( "_" is usually preferred)
* concise as will be comprehensible to others

Try using the `%>%` operator as well. So you will be "piping" `d %>% rename(...)`.

```{r rename}
d.tidy <- ...
```

> ANSWER-START

```{r rename}
d.tidy <- d.tidy %>%
  rename(workerid = WorkerId,
         condition = Input.condition,          
         dog = Answer.dog_cost,
         plasma = Answer.plasma_cost, 
         sushi = Answer.sushi_cost)

```
> ANSWER-END

OK, now for the tricky part. Use the verb *gather* to turn this into a *tidy* data frame.

HINT: look for online examples!

```{r gather}
d.tidy <- ...
```

> ANSWER-START

```{r gather}
d.tidy <- d.tidy %>%
  gather(item, bet, dog, plasma, sushi)
```
> ANSWER-END

Bonus problem: *spread* these data back into a wide format data frame.

```{r spread}
d.wide <- ...
```

> ANSWER-START

```{r spread}
d.wide <- d.tidy %>%
  spread(item, bet)
```
> ANSWER-END

Part 3: Manipulating the data using dplyr
-----------------------------------------

NOTE: If you generally use `plyr` package, note that they do not play nicely together so things like the rename function won't work unless you load `dplyr` after `plyr`.

As we said in class, a good thing to do is always to check histograms of the response variable. Do that now, using either regular base graphics or ggplot. What can you conclude? 

```{r}
...
```

> ANSWER-START

```{r}
ggplot(d.tidy, aes(x=bet))  + 
  geom_histogram() +
  facet_grid(. ~ item)
```
> ANSWER-END

Try also using the dplyr `distinct` function to remove the duplicate participants from the raw csv file that you discovered in part 1.

```{r}
d.raw <- read.csv("data/janiszewski_rep_exercise.csv")
d.unique.subs <- ...
```

> ANSWER-START

```{r}
d.unique.subs <- distinct(d.raw, WorkerId)
```
> ANSWER-END

OK, now we turn to the actual data anlysis. We'll be using dplyr verbs to *filter*, *group*,*mutate*, and *summarise* the data.

Start by using *summarise* to compute the grand mean bet. (Note that this is the same as taking the grant mean - the value will come later. Right now we're just learning the syntax of that verb.)

```{r}
...
```

> ANSWER-START

```{r}
d.tidy %>% summarise(bet = mean(bet, na.rm=TRUE))
```
> ANSWER-END

This is a great time to get comfortable with the %>% operator. In brief, %>% allows you to pipe data from one function to another. So if you would have written:

`d <- function(d, other_stuff)`
you can now write:

`d <- d %>% function(other_stufF)`

That doesn't seem like much, but it's cool when you can replace:

`d <- function1(d, other_stuff)`
`d <- function2(d, lots_of_other_stuff, more_stuff)`
`d <- function3(d, yet_more_stuff)`

with

`d <- d %>%` 
`function1(other_stuff) %>%`
`function2(lots_of_other_stuff, more_stuff) %>%`
`function3(yet_more_stuff)`

In other words, you get to make a clean list of the things you want to do and chain them together without a lot of intermediate assignments. 

Let's use that capacity to combine summarise with *group_by*, which allows us to break up our summary into groups. Try grouping by item and condition and taking means using *summarise*, chaining these two verbs with %>%.

```{r}
...
```

> ANSWER-START

```{r}
d.tidy %>% 
  group_by(item, condition) %>%
  summarise(bet = mean(bet, na.rm=TRUE))
```
> ANSWER-END


OK, it's looking like there are maybe some differences between conditions, but how are we going to plot these? They are fundamentally different magnitudes from one another. 

Really we need the size of the deviation from the anchor, which means we need the anchor value. Let's go back to the data and add that in.

Take a look at this complex expression. You don't have to modify it, but see what is being done here with gather, separate and spread. Run each part (e.g. the first verb, the first two verbs, etc.) and after doing each, look at head(d.tidy) to see what they do.  

```{r}
d.tidy <- d %>% 
  select(WorkerId, Input.condition, 
         starts_with("Answer"), 
         starts_with("Input")) %>%
  rename(workerid = WorkerId,
         condition = Input.condition,          
         dog_anchor = Input.price1,
         plasma_anchor = Input.price2,
         sushi_anchor = Input.price3,
         dog_cost = Answer.dog_cost,
         plasma_cost = Answer.plasma_cost, 
         sushi_cost = Answer.sushi_cost) %>%
  gather(name, cost, 
         dog_anchor, plasma_anchor, sushi_anchor, 
         dog_cost, plasma_cost, sushi_cost) %>%
  separate(name, c("item", "type"), "_") %>%
  spread(type, cost) 
```

Now we can do the same thing as before but look at the relative difference between anchor and estimate. Let's do this two ways: 

* By computing absolute percentage change in price, and 
* By computing z-scores over items.

To do the first, use the *mutate* verb to add a percent change column, then comute the same summary as before. 

```{r}
pcts <- d.tidy %>% 
  mutate(pct_change = ... ) %>% 
  ...
```

> ANSWER-START

```{r}
pcts <- d.tidy %>% 
  mutate(pct_change = abs((anchor - cost) / anchor)) %>%
  group_by(item, condition) %>%
  summarise(pct = mean(pct_change, na.rm=TRUE))
```
> ANSWER-END

To do the second, you will need to *group* once by item, then to *ungroup* and  do the same thing as before. NOTE: you can use group_by(..., add=FALSE) to set new grouping levels, also. 

HINT: `scale(x)` returns a complicated data structure that doesn't play nicely with dplyr. try `scale(x)[,1]` to get what you need. 

```{r}
pcts <- d.tidy %>% 
  mutate(pct_change = ... ) %>% 
  ...
```

> ANSWER-START

```{r}
z.scores <- d.tidy %>% 
  group_by(item) %>%
  mutate(z = scale(anchor - cost)[,1]) %>%
  group_by(item, condition, add=FALSE) %>%
  summarise(z = mean(z, na.rm=TRUE))
```
> ANSWER-END

OK, now here comes the end: we're going to plot the differences and see if anything happened. First the percent change:

```{r}
ggplot(pcts, aes(x=item, y=pct, fill=condition)) + 
      geom_bar(stat="identity", position="dodge")  
```

and the z-scores:

```{r}
ggplot(z.scores, aes(x=item, y=z, fill=condition)) + 
      geom_bar(position="dodge", stat="identity")  
```

Oh well. This replication didn't seem to work out straightforwardly. 

END 
===